---
title: "Model selection and overfitting"
subtitle: "Lecture 18"
date: "March 26, 2024"
format: 
  revealjs:
    footer: "[🔗 sta199-s24.github.io](https://sta199-s24.github.io/) &nbsp;·&nbsp; [❓ Ask on Ed](https://edstem.org/us/courses/50730)"
---

# Warm up

```{r}
#| echo: false
#| message: false
library(countdown)
library(tidyverse)
ggplot2::theme_set(theme_gray(base_size = 16))
```

## While you wait for class to begin...

::: nonincremental
Any questions from prepare materials?
:::

```{=html}
<iframe allowfullscreen frameborder="0" height="100%" mozallowfullscreen style="min-width: 500px; min-height: 355px" src="https://app.wooclap.com/STA199S24?from=status-bar?" width="100%"></iframe>
```
## Announcements

-   Read the feedback (including written feedback) on labs

-   Project repos to be released back to you on Friday

-   Status update emails in your inbox

# Questions from last time {.smaller}

## Can you review the `augment()` function? {.smaller}

The `augment()` function can be used to "augment" a data set (usually of new observations) with the model.

::: incremental
-   It calculates predictions, $\hat{y}$s, under the given model.

-   It also calculates the residuals, $e$s, for each observation.

-   It returns a data frame of the input data *augmented* with predicted values and residuals.
:::

## `augment()` - Setup

```{r}
#| message: false

library(tidyverse)
library(tidymodels)

mtcars_fit <- linear_reg() |>
  fit(mpg ~ disp + cyl, data = mtcars)

tidy(mtcars_fit)
```

## `augment()` - Augment the original data {.smaller}

```{r}
augment(mtcars_fit, new_data = mtcars)
```

## `augment()` - Augment new data

```{r}
new_cars <- tibble(
  mpg = c(23, 22), disp = c(150, 350), cyl = c(4, 6)
)

new_cars
```

. . .

```{r}
augment(mtcars_fit, new_data = new_cars)
```

## What is the difference between $R^2$ and adjusted $R^2$? {.smaller}

-   $R^2$:

    -   Proportion of variability in the outcome explained by the model.

    -   Useful for quantifying the fit of a given model.

. . .

-   Adjusted $R^2$:

    -   Proportion of variability in the outcome explained by the model, with a penalty added for the number of predictors in the model.

    -   Useful for comparing models.

# From last time

## Application exercise: `ae-13-modeling-loans` {.smaller}

::: appex
-   Go to your project called `ae`.
-   Continue working on `ae-13-modeling-loans.qmd`.
:::

**Goals:**

-   Review prediction and interpretation of model results
-   Review main and interaction effects models
-   Discuss model selection further

## Recap: `ae-13-modeling-loans`

-   What is the practical difference between a model with parallel and non-parallel lines?

-   What is the definition of R-squared?

-   Why do we choose models based on adjusted R-squared and not R-squared?
