---
title: "AE 07: Types and classes and populations"
subtitle: "Suggested answers"
categories: 
  - Application exercise
  - Answers
---

::: callout-important
These are suggested answers.
This document should be used as reference only, it's not designed to be an exhaustive key.
:::

## Packages

We will use the following two packages in this application exercise.

-   **tidyverse**: For data import, wrangling, and visualization.
-   **scales**: For better axis labels.

```{r}
#| label: load-packages
#| message: false

library(tidyverse)
library(skimr)
library(scales)
```

## Type coercion

-   **Demo:** Determine the type of the following vector.
    And then, change the type to numeric.

    ```{r}
    #| label: typeof-x

    x <- c("1", "2", "3")
    typeof(x)

    as.numeric(x)
    ```

-   **Demo:** Once again, determine the type of the following vector.
    And then, change the type to numeric.
    What's different than the previous exercise?

    ```{r}
    #| label: typeof-y

    y <- c("a", "b", "c")
    typeof(y)

    as.numeric(y)
    ```

-   **Demo:** Once again, determine the type of the following vector.
    And then, change the type to numeric.
    What's different than the previous exercise?

    ```{r}
    #| label: typeof-z

    z <- c("1", "2", "three")
    typeof(z)

    as.numeric(z)
    ```

-   **Demo:** Suppose you conducted a survey where you asked people how many cars their household owns collectively.
    And the answers are as follows:

    ```{r}
    #| label: survey-results

    survey_results <- tibble(cars = c(1, 2, "three"))
    survey_results
    ```

    This is annoying because of that third survey taker who just had to go and type out the number instead of providing as a numeric value.
    So now you need to update the `cars` variable to be numeric.
    You do the following

    ```{r}
    #| label: survey-results-mutate-1

    survey_results |>
      mutate(cars = as.numeric(cars))
    ```

    And now things are even more annoying because you get a warning `NAs introduced by coercion` that happened while computing `cars = as.numeric(cars)` and the response from the third survey taker is now an `NA` (you lost their data).
    Fix your `mutate()` call to avoid this warning.

    ```{r}
    #| label: survey-results-mutate-2

    survey_results |>
      mutate(
        cars = if_else(cars == "three", "3", cars),
        cars = as.numeric(cars)
      )
    ```

-   **Your turn:** First, guess the type of the vector.
    Then, check if you guessed right.
    I've done the first one for you, you'll see that it's helpful to check the type of each element of the vector first.

    -   `c(1, 1L, "C")`

        ```{r}
        v1 <- c(1, 1L, "C")

        # to help you guess
        typeof(1)
        typeof(1L)
        typeof("C")

        # to check after you guess
        typeof(v1)
        ```

    -   `c(1L / 0, "A")`

        ```{r}
        v2 <- c(1L / 0, "A")

        # to help you guess
        typeof(1L)
        typeof(0)
        typeof(1L / 0)
        typeof("A")

        # to check after you guess
        typeof(v2)
        ```

    -   `c(1:3, 5)`

        ```{r}
        v3 <- c(1:3, 5)

        # to help you guess
        typeof(1:3)
        typeof(5)

        # to check after you guess
        typeof(v3)
        ```

    -   `c(3, "3+")`

        ```{r}
        v4 <- c(3, "3+")

        # to help you guess
        typeof(3)
        typeof("3+")

        # to check after you guess
        typeof(v4)
        ```

    -   `c(NA, TRUE)`

        ```{r}
        v5 <- c(NA, TRUE)

        # to help you guess
        typeof(NA)
        typeof(TRUE)

        # to check after you guess
        typeof(v5)
        ```

## Populations in continents

In the previous application exercise you joined two datasets (after a bit of data cleaning), and calculated total population in each continent and visualized it.

-   First, you loaded the data:

```{r}
#| label: load-data

continents <- read_csv("https://sta199-s24.github.io/data/continents.csv")
population <- read_csv("https://sta199-s24.github.io/data/world-pop-2022.csv")
```

-   Then you cleaned the country names where the spelling in one data frame didn't match the other, and joined the data sets:

```{r}
#| label: clean and join

population_continent <- population |>
  mutate(country = case_when(
    country == "Congo, Dem. Rep." ~ "Democratic Republic of Congo",
    country == "Congo, Rep." ~ "Congo",
    country == "Hong Kong SAR, China" ~ "Hong Kong",
    country == "Korea, Dem. People's Rep." ~ "North Korea",
    country == "Korea, Rep." ~ "South Korea",
    country == "Kyrgyz Republic" ~ "Kyrgyzstan",
    .default = country
    )
  ) |>
  left_join(continents, by = join_by(country == entity))
```

-   Then, you calculated total population for each continent.

```{r}
#| label: summarize

population_summary <- population_continent |>
  group_by(continent) |>
  summarize(total_pop = sum(population)) |>
  arrange(desc(total_pop))
```

-   And finally, you visualized these data.

```{r}
#| label: visualize

ggplot(population_summary) +
  geom_point(aes(x = total_pop, y = continent)) +
  geom_segment(aes(y = continent, yend = continent, x = 0, xend = total_pop)) +
  scale_x_continuous(labels = label_number(scale = 1/1000000, suffix = " bil")) +
  theme_minimal() +
  labs(
    x = "Total population",
    y = "Continent",
    title = "World population",
    subtitle = "As of 2022",
    caption = "Data sources: The World Bank and Our World in Data"
  )
```

-   **Question:** Take a look at the visualization. How are the continents ordered? What would be a better order?

Ordering the continents by the value of total population would be better.

-   **Demo:** Reorder the continents on the y-axis (levels of `continent`) in order of value of total population. You will want to use a function from the **forcats** package, see <https://forcats.tidyverse.org/reference/index.html> for inspiration and help.

```{r}
#| label: visualize-better

population_summary |>
  mutate(continent = fct_reorder(continent, total_pop)) |>
  ggplot() +
  geom_point(aes(x = total_pop, y = continent)) +
  geom_segment(aes(y = continent, yend = continent, x = 0, xend = total_pop)) +
  geom_segment(aes(y = continent, yend = continent, x = 0, xend = total_pop)) +
  scale_x_continuous(labels = label_number(scale = 1/1000000, suffix = " bil")) +
  theme_minimal() +
  labs(
    x = "Total population",
    y = "Continent",
    title = "World population",
    subtitle = "As of 2022",
    caption = "Data sources: The World Bank and Our World in Data"
  )
```

-   **Think out loud:** Describe what is happening in the each step of the code chunk above.

Answers may vary.
