---
title: "AE 09: Opinion articles in The Chronicle"
---

## Part 1 - Data scraping

This will be done in the `chronicle-scrape.R` R script.
Save the resulting data frame in the `data` folder.

## Part 2 - Data analysis

Let's start by loading the packages we will need:

```{r}
#| label: load-packages
#| message: false

library(tidyverse)
```

-   **Your turn (1 minute):** Load the data you saved into the `data` folder and name it `chronicle`.

```{r}
#| label: load-data
#| message: false

# add code here
```

-   **Your turn (3 minutes):** Who are the most prolific authors of the 100 most recent opinion articles in The Chronicle?

```{r}
#| label: union-title-abstract

# add code here
```

-   **Demo:** Draw a line plot of the number of opinion articles published per day in The Chronicle.

```{r}
#| label: articles-per-day

# add code here
```

-   **Demo:** What percent of the most recent 100 opinion articles in The Chronicle mention "climate" in their title?

```{r}
#| label: climate-title

# add code here
```

-   **Your turn (5 minutes):** What percent of the most recent 100 opinion articles in The Chronicle mention "climate" in their title or abstract?

```{r}
#| label: climate-title-abstract

# add code here
```

-   **Time permitting:** Come up with another question and try to answer it using the data.

```{r}
# add code here
```
