---
title: "AE 11: Modelling penguins"
categories: 
  - Application exercise
---

In this application exercise we will be studying penguins.
The data can be found in the **palmerpenguins** package and we will use **tidyverse** and **tidymodels** for data exploration and modeling, respectively.

```{r}
#| label: load-packages
#| message: false

library(tidyverse)
library(tidymodels)
library(palmerpenguins)
```

Please read the following context and take a `glimpse` at the data set before we get started.

> This data set comprising various measurements of three different penguin species, namely Adelie, Gentoo, and Chinstrap.
> The rigorous study was conducted in the islands of the Palmer Archipelago, Antarctica.
> These data were collected from 2007 to 2009 by Dr. Kristen Gorman with the Palmer Station Long Term Ecological Research Program, part of the US Long Term Ecological Research Network.
> The data set is called `penguins`.

```{r}
#| label: glimpse-penguins

# add code here
```

Our goal is to understand better how various body measurements and attributes of penguins relate to their body mass.
First, we are going to investigate the relationship between a penguins' flipper lengths and their body masses.

-   **Question:** Based on our research focus, which variable is the response variable?

*Add response here*

-   **Demo:** Visualize the relationship between flipper length and body mass of penguins.

```{r}
#| label: plot-bm-fl

# add code here 
```

# Correlation

-   **Your turn (5 minutes):**
    -   What is correlation?
        What values can correlation take?

        *Strength and direction of a linear relationship. It's bounded by -1 and 1.*

    -   Are you good at guessing correlation?
        Give it a try!
        <https://www.rossmanchance.com/applets/2021/guesscorrelation/GuessCorrelation.html>
-   **Demo:** What is the correlation between flipper length and body mass of penguins?

```{r}
#| label: correlation

# add code here
```

# Defining, fitting, and summarizing a model

-   **Demo:** Write the population model below that explains the relationship between body mass and flipper length.

$$
add~math~text~here
$$

-   **Demo:** Fit the linear regression model and display the results. Write the estimated model output below.

```{r}
#| label: fit-bm-fl

# add code here
```

$$
add~math~text~here
$$

-   **Your turn:** Interpret the slope and the intercept in the context of the data.

    -   Intercept: *Add your response here*

    -   Slopes: *Add your response here*

-   **Your turn:** Recreate the visualization from above, this time adding a regression line to the visualization `geom_smooth(method = "lm")`.

```{r}
#| label: plot-bm-fl-line

# add code here
```

-   What is the estimated body mass for a penguin with a flipper length of 210?

```{r}
#| label: predict-210

# add code here
```

-   What is the estimated body mass for a penguin with a flipper length of 100?

```{r}
#| label: predict-100

# add code here
```

# Another model

-   **Demo:** A different researcher wants to look at body weight of penguins based on the island they were recorded on. How are the variables involved in this analysis different?

*Add response here*

-   **Demo:** Make an appropriate visualization to investigate this relationship below. Additionally, calculate the mean body mass by island.

```{r}
#| label: plot-bm-island

# add code here
```

```{r}
#| label: mean-bm-island

# add code here
```

-   **Demo:** Change the geom of your previous plot to `geom_point()`. Use this plot to think about how R models these data.

```{r}
#| label: plot-bm-island-points

# add code here
```

-   **Your turn:** Fit the linear regression model and display the results. Write the estimated model output below.

```{r}
#| label: fit-bm-island

# add code here
```

Add math text here

-   **Demo:** Interpret each coefficient in context of the problem.

    -   Intercept: *Add your response here*

    -   Slopes: *Add your response here*

-   **Demo:** What is the estimated body weight of a penguin on Biscoe island?
    What are the estimated body weights of penguins on Dream and Torgersen islands?

```{r}
#| label: predict-bm-island

# add code here
```
