---
title: "AE 10: Modelling fish"
subtitle: "Suggested answers"
categories: 
  - Application exercise
  - Answers
fig-width: 7
fig-asp: 0.618
---

::: callout-important
These are suggested answers.
This document should be used as reference only, it's not designed to be an exhaustive key.
:::

For this application exercise, we will work with data on fish.
The dataset we will use, called `fish`, is on two common fish species in fish market sales.

```{r}
#| label: load-packages-data
#| message: false

library(tidyverse)
library(tidymodels)

fish <- read_csv("data/fish.csv")
```

The data dictionary is below:

| **variable**      | **description**        |
|-------------------|------------------------|
| `species`         | Species name of fish   |
| `weight`          | Weight, in grams       |
| `length_vertical` | Vertical length, in cm |
| `length_diagonal` | Diagonal length, in cm |
| `length_cross`    | Cross length, in cm    |
| `height`          | Height, in cm          |
| `width`           | Diagonal width, in cm  |

# Visualizing the model

We're going to investigate the relationship between the weights and heights of fish.

-   **Demo:** Create an appropriate plot to investigate this relationship. Add appropriate labels to the plot.

```{r}
#| label: plot-hw

ggplot(fish, aes(x = height, y = weight)) +
  geom_point() +
  labs(
    title = "Weights vs. heights of fish",
    x = "Height (cm)",
    y = "Weight (gr)"
  )
```

-   **Your turn (5 minutes):**

    -   If you were to draw a a straight line to best represent the relationship between the heights and weights of fish, where would it go?
        Why?

        *Start from the bottom and go up Identify the first and last point and draw a line through most the others.*

    -   Now, let R draw the line for you.
        Refer to the documentation at <https://ggplot2.tidyverse.org/reference/geom_smooth.html>.
        Specifically, refer to the `method` section.

    ```{r}
    #| label: plot-hw-lm

    ggplot(fish, aes(x = height, y = weight)) +
      geom_point() +
      geom_smooth(method = "lm") +
      labs(
        title = "Weights vs. lengths of fish",
        x = "Head-to-tail lentgh (cm)",
        y = "Weight of fish (grams)"
      )
    ```

    -   What types of questions can this plot help answer?

        *Is there a relationship between fish heights and weights of fish?*

-   ***Your turn (3 minutes):***

    -   We can use this line to make predictions.
        Predict what you think the weight of a fish would be with a height of 10 cm, 15 cm, and 20 cm.
        Which prediction is considered extrapolation?

        *At 10 cm, we estimate a weight of 375 grams. At 15 cm, we estimate a weight of 600 grams At 20 cm, we estimate a weight of 975 grams. 20 cm would be considered extrapolation.*

    -   What is a residual?

        *Difference between predicted and observed.*

# Model fitting

-   **Demo:** Fit a model to predict fish weights from their heights.

```{r}
#| label: fit-model

fish_hw_fit <- linear_reg() |>
  fit(weight ~ height, data = fish)

fish_hw_fit
```

-   **Your turn (3 minutes):** Predict what the weight of a fish would be with a height of 10 cm, 15 cm, and 20 cm using this model.

```{r}
#| label: predict-by-hand

x <- c(10, 15, 20)
-288 + 60.92 * x
```

-   **Demo:** Calculate predicted weights for all fish in the data and visualize the residuals under this model.

```{r}
#| label: residuals

fish_hw_aug <- augment(fish_hw_fit, new_data = fish)

ggplot(fish_hw_aug, aes(x = height, y = weight)) + 
  geom_point() + 
  geom_smooth(method = "lm", se = FALSE, color = "lightgrey") +  
  geom_segment(aes(xend = height, yend = .pred), color = "gray") +  
  geom_point(aes(y = .pred), shape = "circle open") + 
  theme_minimal() +
  labs(
    title = "Weights vs. heights of fish",
    subtitle = "Residuals",
    x = "Height (cm)",
    y = "Weight (gr)"
  )
```

# Model summary

-   **Demo:** Display the model summary including estimates for the slope and intercept along with measurements of uncertainty around them. Show how you can extract these values from the model output.

```{r}
#| label: model-summary

fish_hw_tidy <- tidy(fish_hw_fit)
fish_hw_tidy
```

-   **Demo:** Write out your model using mathematical notation.

$\widehat{weight} = -288 + 60.9 \times height$

# Correlation

We can also assess correlation between two quantitative variables.

-   **Your turn (5 minutes):**
    -   What is correlation?
        What are values correlation can take?

        *Strength and direction of a linear relationship. It's bounded by -1 and 1.*

    -   Are you good at guessing correlation?
        Give it a try!
        <https://www.rossmanchance.com/applets/2021/guesscorrelation/GuessCorrelation.html>
-   **Demo:** What is the correlation between heights and weights of fish?

```{r}
fish |>
  summarize(r = cor(height, weight))
```

# Adding a third variable

-   **Demo:** Does the relationship between heights and weights of fish change if we take into consideration species? Plot two separate straight lines for the Bream and Roach species.

```{r}
#| label: plot-hws-lm

ggplot(fish, 
       aes(x = height, y = weight, color = species)) +
  geom_point() +
  geom_smooth(method = "lm") +
  labs(
    title = "Weights vs. heights of fish",
    x = "Height (cm)",
    y = "Weight (gr)"
  )
```

# Fitting other models

-   **Demo:** We can fit more models than just a straight line. Change the following code below to read `method = "loess"`. What is different from the plot created before?

```{r}
#| label: plot-hw-loess

ggplot(fish, 
       aes(x = height, y = weight)) +
  geom_point() +
  geom_smooth(method = "lm") +
  labs(
    title = "Weights vs. heights of fish",
    x = "Height (cm)",
    y = "Weight (gr)"
  )
```
