---
title: "Modelling fish"
---

For this application exercise, we will work with data on fish.
The dataset we will use, called `fish`, is on two common fish species in fish market sales.

```{r}
#| label: load-packages-data
#| message: false

library(tidyverse)
library(tidymodels)

fish <- read_csv("data/fish.csv")
```

The data dictionary is below:

| **variable**      | **description**        |
|-------------------|------------------------|
| `species`         | Species name of fish   |
| `weight`          | Weight, in grams       |
| `length_vertical` | Vertical length, in cm |
| `length_diagonal` | Diagonal length, in cm |
| `length_cross`    | Cross length, in cm    |
| `height`          | Height, in cm          |
| `width`           | Diagonal width, in cm  |

# Visualizing the model

We're going to investigate the relationship between the weights and heights of fish.

-   **Demo:** Create an appropriate plot to investigate this relationship. Add appropriate labels to the plot.

```{r}
#| label: plot-hw

# add code here
```

-   **Your turn (5 minutes):**

    -   If you were to draw a a straight line to best represent the relationship between the heights and weights of fish, where would it go?
        Why?

        *Add response here.*

    -   Now, let R draw the line for you.
        Refer to the documentation at <https://ggplot2.tidyverse.org/reference/geom_smooth.html>.
        Specifically, refer to the `method` section.

    ```{r}
    #| label: plot-hw-lm

    # add code here
    ```

    -   What types of questions can this plot help answer?

    *Add response here.*

-   **Your turn (3 minutes):**

    -   We can use this line to make predictions. Predict what you think the weight of a fish would be with a height of 10 cm, 15 cm, and 20 cm. Which prediction is considered extrapolation?

    *Add response here.*

    -   What is a residual?

    *Add response here.*

# Model fitting

-   **Demo:** Fit a model to predict fish weights from their heights.

```{r}
#| label: fit-model

# add code here
```

-   **Your turn (3 minutes):** Predict what the weight of a fish would be with a height of 10 cm, 15 cm, and 20 cm using this model.

```{r}
#| label: predict-by-hand

# add code here
```

-   **Demo:** Calculate predicted weights for all fish in the data and visualize the residuals under this model.

```{r}
#| label: residuals

# add code here
```

# Model summary

-   **Demo:** Display the model summary including estimates for the slope and intercept along with measurements of uncertainty around them. Show how you can extract these values from the model output.

```{r}
#| label: model-summary

# add code here
```

-   **Demo:** Write out your model using mathematical notation.

*Add response here.*

# Correlation

We can also assess correlation between two quantitative variables.

-   **Your turn (5 minutes):**

    -   What is correlation? What are values correlation can take?

    *Add response here.*

    -   Are you good at guessing correlation? Give it a try! <https://www.rossmanchance.com/applets/2021/guesscorrelation/GuessCorrelation.html>

-   **Demo:** What is the correlation between heights and weights of fish?

```{r}
#| label: correlation

# add code here
```

# Adding a third variable

-   **Demo:** Does the relationship between heights and weights of fish change if we take into consideration species? Plot two separate straight lines for the Bream and Roach species.

```{r}
#| label: plot-hws-lm

# add code here
```

# Fitting other models

-   **Demo:** We can fit more models than just a straight line. Change the following code below to read `method = "loess"`. What is different from the plot created before?

```{r}
#| label: plot-hw-loess

# add code here
```
