---
title: "AE 15: Modeling houses in Duke Forest"
categories: 
  - Application exercise
---

In this application exercise, we will

-   use bootstrapping to quantify the uncertainty around a measure of center – median
-   use bootstrapping to quantify the uncertainty around a measure of relationship – slope
-   interpret confidence intervals

The dataset are on housing prices in Duke Forest – a dataset you've seen before!
It's called `duke_forest` and it's in the **openintro** package.
Additionally, we'll use **tidyverse** and **tidymodels** packages.

```{r}
#| label: load-packages
#| message: false
library(tidyverse)
library(tidymodels)
library(openintro)
```

# Typical size of a house in Duke Forest

## Exercise 1

Visualize the distribution of sizes of houses in Duke Forest.
What is the size of a typical house?

```{r}
#| label: df-area-hist
# add code here
```

## Exercise 2

Construct a 95% confidence interval for the typical size of a house in Duke Forest.
Interpret the interval in context of the data.

```{r}
#| label: df-area-median-boot
# add code here
```

*Add interpretation here.*

## Exercise 3

Without calculating it – would a 90% confidence interval be wider or narrower?
Why?

*Add response here.*

## Exercise 4

Construct the 90% confidence interval and interpret it.

```{r}
#| label: df-area-median-ci
# add code here
```

*Add interpretation here.*

# Relationship between price and size

The following model predicts price of a house in Duke Forest from its size.

```{r}
#| label: df-price-area-fit
df_price_area_fit <- linear_reg() |>
  fit(price ~ area, data = duke_forest)

tidy(df_price_area_fit)
```

The slope can be interpreted as:

> For each additional square feet, the model predicts that prices of houses in Duke Forest are higher by \$159, on average.

## Exercise 5

Quantify the uncertainty around this slope using a 95% bootstrap confidence interval and interpret the interval in context of the data.

```{r}
#| label: df-price-area-boot
# add code here
```

*Add interpretation here.*
