---
title: "Modelling loan interest rates"
subtitle: "Suggested answers"
categories: 
  - Application exercise
  - Answers
editor: visual
---

In this application exercise we will be studying loan interest rates.
The dataset is one you've come across before in your reading -- the dataset about loans from the peer-to-peer lender, Lending Club, from the **openintro** package.
We will use **tidyverse** and **tidymodels** for data exploration and modeling, respectively.

```{r}
#| label: load-packages
#| message: false

library(tidyverse)
library(tidymodels)
library(openintro)
```

Before we use the dataset, we'll make a few transformations to it.

-   **Your turn:** Review the code below with your neighbor and write a summary of the data transformation pipeline.

*Add response here.*

```{r}
#| label: data-prep

loans <- loans_full_schema |>
  mutate(
    credit_util = total_credit_utilized / total_credit_limit,
    bankruptcy = as.factor(if_else(public_record_bankrupt == 0, 0, 1)),
    verified_income = droplevels(verified_income),
    homeownership = str_to_title(homeownership),
    homeownership = fct_relevel(homeownership, "Rent", "Mortgage", "Own")
  ) |>
  rename(credit_checks = inquiries_last_12m) |>
  select(
    interest_rate, loan_amount, verified_income, 
    debt_to_income, credit_util, bankruptcy, term, 
    credit_checks, issue_month, homeownership
  )
```

Here is a glimpse at the data:

```{r}
#| label: glimpse-loans

glimpse(loans)
```

# Get to know the data

-   **Your turn:** What is a typical interest rate in this dataset?
    What are some attributes of a typical loan and a typical borrower.
    Give yourself no more than 5 minutes for this exploration and share 1-2 findings.

```{r}
#| label: explore-loans
#| layout-ncol: 2
#| warning: false

ggplot(loans, aes(x = interest_rate)) +
  geom_histogram(binwidth = 1)

ggplot(loans, aes(x = loan_amount)) +
  geom_histogram(binwidth = 5000)

ggplot(loans, aes(x = term)) +
  geom_bar()

ggplot(loans, aes(x = issue_month)) +
  geom_bar()
```

```{r}
#| label: explore-borrowers
#| layout-ncol: 2
#| warning: false

ggplot(loans, aes(x = credit_util)) +
  geom_histogram(binwidth = 0.1)

ggplot(loans, aes(x = verified_income)) +
  geom_bar()

ggplot(loans, aes(x = debt_to_income)) +
  geom_histogram(binwidth = 10)

ggplot(loans, aes(x = bankruptcy)) +
  geom_bar()

ggplot(loans, aes(x = credit_checks)) +
  geom_bar()

ggplot(loans, aes(x = homeownership)) +
  geom_bar()
```

# Interest rate vs. credit utilization ratio

The regression model for interest rate vs. credit utilization is as follows.

```{r}
#| label: rate-util-fit

rate_util_fit <- linear_reg() |>
  fit(interest_rate ~ credit_util, data = loans)

tidy(rate_util_fit)
```

And here is the model visualized:

```{r}
#| label: rate-util-plot
#| warning: false

ggplot(loans, aes(x = credit_util, y = interest_rate)) +
  geom_point(alpha = 0.5) +
  geom_smooth(method = "lm")
```

-   **Your turn:** Interpret the intercept and the slope.

Intercept: Borrowers with 0 credit utilization are predicted, on average, to get an interest rate of 10.5%.

Slope: For each additional point credit utilization is higher, interest rate is predicted to be higher, on average, by 4.73%.

# Interest rate vs. homeownership

Next we predict interest rates from homeownership, which is a categorical predictor with three levels:

```{r}
levels(loans$homeownership)
```

-   **Demo:** Fit the linear regression model to predict interest rate from homeownership and display a tidy summary of the model. Write the estimated model output below.

```{r}
#| label: rate-home-fit

rate_home_fit <- linear_reg() |>
  fit(interest_rate ~ homeownership, data = loans)

tidy(rate_home_fit)
```

-   **Your turn:** Interpret each coefficient in context of the problem.

    -   Intercept: Loan applicants who rent are predicted to receive an interest rate of 12.9%, on average.

    -   Slopes:

        -   The model predicts that loan applicants who have a mortgage for their home receive 0.866% lower interest rate than those who rent their home, on average.

        -   The model predicts that loan applicants who own their home receive 0.611% lower interest rate than those who rent their home, on average.

# Interest rate vs. credit utilization and homeownership

## Main effects model

-   **Demo:** Fit a model to predict interest rate from credit utilization and homeownership, **without** an interaction effect between the two predictors. Display the summary output and write out the estimated regression equation.

```{r}
#| label: rate-util-home-fit

rate_util_home_fit <- linear_reg() |>
  fit(interest_rate ~ credit_util + homeownership, data = loans)

tidy(rate_util_home_fit)
```

$$
\widehat{interest~rate} = 9.93 + 5.34 \times credit~util + 0.696 \times Mortgage + 0.128 \times Own
$$

-   **Demo:** Write the estimated regression equation for loan applications from each of the homeownership groups separately.
    -   Rent: $\widehat{interest~rate} = 9.93 + 5.34 \times credit~util$
    -   Mortgage: $\widehat{interest~rate} = 10.626 + 5.34 \times credit~util$
    -   Own: $\widehat{interest~rate} = 10.058 + 5.34 \times credit~util$
-   **Question:** How does the model predict the interest rate to vary as credit utilization varies for loan applicants with different homeownership status. Are the rates the same or different?

The same.

## Interaction effects model

-   **Demo:** Fit a model to predict interest rate from credit utilization and homeownership, **with** an interaction effect between the two predictors. Display the summary output and write out the estimated regression equation.

```{r}
#| label: rate-util-home-int-fit

rate_util_home_int_fit <- linear_reg() |>
  fit(interest_rate ~ credit_util * homeownership, data = loans)

tidy(rate_util_home_int_fit)
```

$$
\widehat{interest~rate} = 9.44 + 6.20 \times credit~util + 1.39 \times Mortgage + 0.697 \times Own - 1.64 \times credit_util:Mortgage - 1.06 \times credit_util:Own
$$

-   **Demo:** Write the estimated regression equation for loan applications from each of the homeownership groups separately.
    -   Rent: $\widehat{interest~rate} = 9.44 + 6.20 \times credit~util$
    -   Mortgage: $\widehat{interest~rate} = 10.83 + 4.56 \times credit~util$
    -   Own: $\widehat{interest~rate} = 10.137 + 5.14 \times credit~util$
-   **Question:** How does the model predict the interest rate to vary as credit utilization varies for loan applicants with different homeownership status. Are the rates the same or different?

Different.

## Choosing a model

Rule of thumb: **Occam's Razor** - Don't overcomplicate the situation!
We prefer the *simplest* best model.

```{r}
#| label: glance

glance(rate_util_home_fit)
glance(rate_util_home_int_fit)
```

-   **Review:** What is R-squared? What is adjusted R-squared?

R-squared is the percent variability in the response that is explained by our model.
(Can use when models have same number of variables for model selection)

Adjusted R-squared is similar, but has a penalty for the number of variables in the model.
(Should use for model selection when models have different numbers of variables).

-   **Question:** Based on the adjusted $R^2$s of these two models, which one do we prefer?

The interaction effects model, though just barely.

# Another model to consider

-   **Your turn:** Let's add one more model to the variable -- issue month. Should we add this variable to the interaction effects model from earlier?

```{r}
#| label: new-fit

linear_reg() |>
  fit(interest_rate ~ credit_util * homeownership + issue_month, data = loans) |>
  glance()
```

No, the adjusted R-squared goes down.
