---
title: "AE 11: Modelling penguins"
subtitle: "Suggested answers"
categories: 
  - Application exercise
  - Answers
fig-width: 7
fig-asp: 0.618
---

In this application exercise we will be studying penguins.
The data can be found in the **palmerpenguins** package and we will use **tidyverse** and **tidymodels** for data exploration and modeling, respectively.

```{r}
#| label: load-packages
#| message: false

library(tidyverse)
library(tidymodels)
library(palmerpenguins)
```

Please read the following context and take a `glimpse` at the data set before we get started.

> This data set comprising various measurements of three different penguin species, namely Adelie, Gentoo, and Chinstrap.
> The rigorous study was conducted in the islands of the Palmer Archipelago, Antarctica.
> These data were collected from 2007 to 2009 by Dr. Kristen Gorman with the Palmer Station Long Term Ecological Research Program, part of the US Long Term Ecological Research Network.
> The data set is called `penguins`.

```{r}
#| label: glimpse-penguins

glimpse(penguins)
```

Our goal is to understand better how various body measurements and attributes of penguins relate to their body mass.
First, we are going to investigate the relationship between a penguins' flipper lengths and their body masses.

-   **Question:** Based on our research focus, which variable is the response variable?

Body mass.

-   **Demo:** Visualize the relationship between flipper length and body mass of penguins.

```{r}
#| label: plot-bm-fl

ggplot(penguins, aes(x = flipper_length_mm, y = body_mass_g)) +
  geom_point()
```

# Correlation

-   **Your turn (5 minutes):**
    -   What is correlation?
        What values can correlation take?

        *Strength and direction of a linear relationship. It's bounded by -1 and 1.*

    -   Are you good at guessing correlation?
        Give it a try!
        <https://www.rossmanchance.com/applets/2021/guesscorrelation/GuessCorrelation.html>
-   **Demo:** What is the correlation between flipper length and body mass of penguins?

```{r}
#| label: correlation

# option 1
penguins |>
  summarize(r = cor(flipper_length_mm, body_mass_g, use = "complete.obs"))

# option 2
penguins |>
  drop_na(flipper_length_mm, body_mass_g) |>
  summarize(r = cor(flipper_length_mm, body_mass_g))
```

# Defining, fitting, and summarizing a model

-   **Demo:** Write the population model below that explains the relationship between body mass and flipper length.

$$
body~mass = \beta_0 + \beta_1 \times flipper~length + \epsilon
$$

-   **Demo:** Fit the linear regression model and display the results. Write the estimated model output below.

```{r}
#| label: fit-bm-fl

bm_fl_fit <- linear_reg() |>
  fit(body_mass_g ~ flipper_length_mm, data = penguins)

tidy(bm_fl_fit)
```

$$
\widehat{body~mass} = -5781 + 49.7 \times flipper~length
$$

-   **Your turn:** Interpret the slope and the intercept in the context of the data.

    -   Intercept: Penguins with 0 flipper length are expected, on average, to weigh 5,781 grams.

    -   Slopes: For each additional millimeter of a penguin;s flipper length, the weight of their penguin is expected to be higher, on average, by 49.7 grams.

-   **Your turn:** Recreate the visualization from above, this time adding a regression line to the visualization `geom_smooth(method = "lm")`.

```{r}
#| label: plot-bm-fl-line

ggplot(penguins,
       aes(x = flipper_length_mm, y = body_mass_g)) +
  geom_point() +
  geom_smooth(method = "lm")
```

-   **Your turn:** What is the estimated body mass for a penguin with a flipper length of 210?

```{r}
#| label: predict-210

bm_fl_fit |>
  predict(new_data = tibble(flipper_length_mm = 210))
```

-   **Demo:** What is the estimated body mass for a penguin with a flipper length of 100?

But we shouldn't do this prediction based on this model since 100 mm is outside of the range of the data (extrapolation).

```{r}
#| label: predict-100

bm_fl_fit |>
  predict(new_data = tibble(flipper_length_mm = 100))
```

# Another model

-   **Demo:** A different researcher wants to look at body weight of penguins based on the island they were recorded on. How are the variables involved in this analysis different?

Predictor is categorical.

-   **Demo:** Make an appropriate visualization to investigate this relationship below. Additionally, calculate the mean body mass by island.

```{r}
#| label: plot-bm-island

ggplot(penguins, aes(x = island, y = body_mass_g)) +
  geom_boxplot()
```

```{r}
#| label: mean-bm-island

penguins |>
  group_by(island) |>
  summarize(mean_bm = mean(body_mass_g, na.rm = TRUE))
```

-   **Demo:** Change the geom of your previous plot to `geom_point()`. Use this plot to think about how R models these data.

```{r}
#| label: plot-bm-island-points

ggplot(penguins, aes(x = island, y = body_mass_g)) +
  geom_point()
```

-   **Your turn:** Fit the linear regression model and display the results. Write the estimated model output below.

```{r}
#| label: fit-bm-island

bm_island_fit <- linear_reg() |>
  fit(body_mass_g ~ island, data = penguins)

tidy(bm_island_fit)
```

-   **Demo:** Interpret each coefficient in context of the problem.

    -   Intercept: Penguins from Biscoe island are expected to weigh, on average, 4,716 grams.

    -   Slopes:

        -   Penguins from Dream island are expected to weigh, on average, 1,003 grams less than those from Biscoe island.

        -   Penguins from Torgersen island are expected to weigh, on average, 1,010 grams less than those from Biscoe island.

-   **Demo:** What is the estimated body weight of a penguin on Biscoe island?
    What are the estimated body weights of penguins on Dream and Torgersen islands?

```{r}
#| label: predict-bm-island

bm_island_fit |>
  predict(new_data = tibble(island = c("Biscoe", "Dream", "Torgersen")))
```
