---
title: "AE 15: Modeling houses in Duke Forest"
subtitle: "Suggested answers"
categories: 
  - Application exercise
  - Answers
editor_options: 
  chunk_output_type: console
---

In this application exercise, we will

-   use bootstrapping to quantify the uncertainty around a measure of center – median
-   use bootstrapping to quantify the uncertainty around a measure of relationship – slope
-   interpret confidence intervals

The dataset are on housing prices in Duke Forest – a dataset you've seen before!
It's called `duke_forest` and it's in the **openintro** package.
Additionally, we'll use **tidyverse** and **tidymodels** packages.

```{r}
#| label: load-packages
#| message: false
library(tidyverse)
library(tidymodels)
library(openintro)
```

# Typical size of a house in Duke Forest

## Exercise 1

Visualize the distribution of sizes of houses in Duke Forest.
What is the size of a typical house?

```{r}
#| label: df-area-hist
ggplot(duke_forest, aes(x = area)) +
  geom_histogram(binwidth = 250)
```

## Exercise 2

Construct a 95% confidence interval for the typical size of a house in Duke Forest.
Interpret the interval in context of the data.

```{r}
#| label: df-area-median-boot
set.seed(12345)
df_araa_median_boot <- duke_forest |>
  specify(response = area) |>
  generate(reps = 100, type = "bootstrap") |>
  calculate(stat = "median")

ggplot(df_araa_median_boot, aes(x = stat)) +
  geom_histogram(binwidth = 50)

df_araa_median_boot |>
  summarize(
    l = quantile(stat, 0.025),
    u = quantile(stat, 0.975)
  )
```

*We are 95% confident that the median house in Duke Forest is between 2,365 and 2,836 square feet.*

## Exercise 3

Without calculating it – would a 90% confidence interval be wider or narrower?
Why?

*Narrower, lower confidence level needed so we can be more precise.*

## Exercise 4

Construct the 90% confidence interval and interpret it.

```{r}
#| label: df-area-median-ci
df_araa_median_boot |>
  summarize(
    l = quantile(stat, 0.05),
    u = quantile(stat, 0.95)
  )
```

*We are 95% confident that the median house in Duke Forest is between 2,395 and 2,830 square feet.*

# Relationship between price and size

The following model predicts price of a house in Duke Forest from its size.

```{r}
#| label: df-price-area-fit
df_price_area_fit <- linear_reg() |>
  fit(price ~ area, data = duke_forest)

tidy(df_price_area_fit)
```

The slope can be interpreted as:

> For each additional square feet, the model predicts that prices of houses in Duke Forest are higher by \$159, on average.

## Exercise 5

Quantify the uncertainty around this slope using a 95% bootstrap confidence interval and interpret the interval in context of the data.

```{r}
#| label: df-price-area-boot
df_price_area_boot <- duke_forest |>
  specify(price ~ area) |>
  generate(reps = 1000, type = "bootstrap") |>
  calculate(stat = "slope")

ggplot(df_price_area_boot, aes(x = stat)) +
  geom_histogram(binwidth = 10)

df_price_area_boot |>
  summarize(
    l = quantile(stat, 0.025),
    u = quantile(stat, 0.975)
  )
```

*We are 95% confident that, for each additional square feet, the model predicts that prices of houses in Duke Forest are higher by \$89.9 to \$214, on average.*
