---
title: "Lab 2 - Data wrangling"
editor_options: 
  chunk_output_type: console
---

# Introduction

In this lab you'll continue to hone your data science workflow as well as bring together what you learned so far in the course (data visualization) with what's coming up (data wrangling).

::: callout-note
This lab assumes you've completed [Lab 0](/labs/lab-0.html) and [Lab 1](/labs/lab-1.html) and doesn't repeat setup and overview content from those labs.
If you have not yet done those, you should go back and review them before starting with this one.
:::

## Learning objectives

By the end of the lab, you will...

-   Be able to create transform data using `dplyr`
-   Build on your mastery of data visualizations using `ggplot2`
-   Get more experience with data science workflow using R, RStudio, Git, and GitHub
-   Further your reproducible authoring skills with Quarto
-   Improve your familiarity with version control using Git and GitHub

## Getting started

### Log in to RStudio

-   Go to <https://cmgr.oit.duke.edu/containers> and log in with your Duke NetID and Password.
-   Click `STA198-199` under My reservations to log into your container. You should now see the RStudio environment.

### Clone the repo & start new RStudio project

-   Go to the course organization at [github.com/sta199-s24](https://github.com/sta199-s24) organization on GitHub.
    Click on the repo with the prefix **lab-2**.
    It contains the starter documents you need to complete the lab.

-   Click on the green **CODE** button, select **Use SSH** (this might already be selected by default, and if it is, you'll see the text **Clone with SSH**).
    Click on the clipboard icon to copy the repo URL.

-   In RStudio, go to *File* ➛ *New Project* ➛*Version Control* ➛ *Git*.

-   Copy and paste the URL of your assignment repo into the dialog box *Repository URL*. Again, please make sure to have *SSH* highlighted under *Clone* when you copy the address.

-   Click *Create Project*, and the files from your GitHub repo will be displayed in the *Files* pane in RStudio.

-   Click *lab-2.qmd* to open the template Quarto file.
    This is where you will write up your code and narrative for the lab.

### First steps

In `lab-2.qmd`, update the `author` field to your name, render your document and examine the changes.
Then, in the Git pane, click on **Diff** to view your changes, add a commit message (e.g., "Added author name"), and click **Commit**.
Then, push the changes to your GitHub repository, and in your browser confirm that these changes have indeed propagated to your repository.

::: callout-important
If you run into any issues with the first steps outlined above, flag a TA for help before proceeding.
:::

## Packages

In this lab we will work with the **tidyverse** package, which is a collection of packages for doing data analysis in a "tidy" way.

```{r}
#| eval: true
#| message: false

library(tidyverse)
```

**Render** the document which loads this package with the `library()` function.

::: callout-note
Remember: The rendered document will include a message about which packages the tidyverse package is loading along with it.
It's just R being informative, a **message** does not indicate anything is wrong (it's not a **warning** or an **error**).
:::

## Guidelines

As we've discussed in lecture, your plots should include an informative title, axes should be labeled, and careful consideration should be given to aesthetic choices.

In addition, the code should all the code should be be able to be read (not run off the page) when you render to PDF.
Make sure that is the case, and add line breaks where the code is running off the page.[^1]

[^1]: Remember, haikus not novellas when writing code!

::: callout-note
Continuing to develop a sound workflow for reproducible data analysis is important as you complete the lab and other assignments in this course.
There will be periodic reminders in this assignment to remind you to **render, commit, and push** your changes to GitHub.
You should have at least 3 commits with meaningful commit messages by the end of the assignment.
:::

# Questions

## Part 1

**Back to the Midwest!**

In this part you will revisit and build on some of your findings from Lab 1, where you explored the `midwest` data frame.
Remember that this data frame is bundled with the **ggplot2** package and is automatically loaded when you load the tidyverse package.
As a refresher, the data contains demographic characteristics of counties in the Midwest region of the United States.
You can read documentation for the data set, including variable definitions by typing `?midwest` in the Console or searching for `midwest` in the Help pane.

### Question 1

Calculate the number of counties in each state and display your results in descending order of number of counties.
Which state has the highest number of counties, and how many?
Which state has the lowest number, and how many?

::: callout-note
The number of counties in a state can change over time, so the values you see in this output may not be true today.
:::

::: render-commit-push
Render, commit, and push your changes to GitHub with the commit message "Added answer for Question 1".

Make sure to commit and push all changed files so that your Git pane is empty afterward.
:::

### Question 2

While in a given state two counties can't have the same name, across states some county names might be shared.
A classmate says "Look at that, there is a county called \_\_\_ in each state in this dataset!" In a single pipeline, discover all counties that could fill in the blank.
Your response should be a data frame with only the county names that could fill in the blank and how many times they appear in the data.

::: callout-tip
You will want to use the `filter()` function in your answer, which requires a logical condition to describe what you want to filter for.
For example `filter(x > 2)` means filter for values of `x` greater than 2, `filter(y <= 3)` means filter for values of y less than or equal to 3.

The table below is a summary of logical operators and how to articulate them in R.

| operator      | definition                |
|---------------|---------------------------|
| `<`           | less than                 |
| `<=`          | less than or equal to     |
| `>`           | greater than              |
| `>=`          | greater than or equal to  |
| `==`          | exactly equal to          |
| `!=`          | not equal to              |
| `x & y`       | `x` AND `y`               |
| `x` \| `y`    | `x` OR `y`                |
| `is.na(x)`    | test if `x` is `NA`       |
| `!is.na(x)`   | test if `x` is not `NA`   |
| `x %in% y`    | test if `x` is in `y`     |
| `!(x %in% y)` | test if `x` is not in `y` |
| `!x`          | not `x`                   |
:::

::: render-commit-push
Render, commit, and push your changes to GitHub with the commit message "Added answer for Question 2".

Make sure to commit and push all changed files so that your Git pane is empty afterward.
:::

### Question 3

Return to the following box plot of population densities where you were asked to identify at least one outlier.

```{r}
#| label: popdensity-boxplot
#| fig-width: 5
#| fig-height: 3
#| echo: false
#| eval: true

ggplot(midwest, aes(x = popdensity)) +
  geom_boxplot() +
  labs(
    title = "Population densities of midwest counties",
    x = "Population density"
  )
```

In this question part, we want you to revisit this box plot and identify the counties described in each part.

a\.
The counties with a population density higher than 25,000.
Your code must use the `filter()` function.

b\.
The county with the highest population density.
Your code must use the `max()` function.

Answer using a single data wrangling pipeline for each part.
Your response should be a data frame five columns: county name, state name, population density, total population, and area, in this order.
If your response has multiple rows, the data frame should be arranged in descending order of population density.

::: render-commit-push
Render, commit, and push your changes to GitHub with the commit message "Added answer for Question 3".

Make sure to commit and push all changed files so that your Git pane is empty afterward.
:::

### Question 4

In Lab 1 you were also asked to describe the distribution of population densities.
The following is one acceptable description that touches on shape, center, and spread of this distribution.
Calculate the values that should go into the blanks.

> The distribution of population density of counties is unimodal and extremely right-skewed.
> A typical Midwestern county has population density of \_\_\_\_ people per unit area.
> The middle 50% of the counties have population densities between \_\_\_ to \_\_\_ people per unit area.

::: callout-tip
Think about which measures of center and spread are appropriate for skewed distributions.
:::

::: render-commit-push
Render, commit, and push your changes to GitHub with the commit message "Added answer for Question 4".

Make sure to commit and push all changed files so that your Git pane is empty afterward.
:::

### Question 5

Another visualization from Lab 1 was the following that showed the proportion of urban counties in each state.

```{r}
#| label: metro-barplot
#| fig-width: 7
#| fig-height: 3
#| echo: false
#| eval: true

midwest <- midwest |>
  mutate(metro = if_else(inmetro == 1, "Yes", "No"))

ggplot(midwest, aes(x = state, fill = metro)) +
  geom_bar(position = "fill") +
  labs(
    title = "Do some states have more urban counties than others?",
    x = "State",
    y = "Proportion in metro area",
    fill = "In metro area?"
  )
```

Calculate these proportions in a single data pipeline.

::: callout-tip
Remember, you'll first need to create a new variable called `metro` which takes on the value `Yes` if the value of `inmetro` is 1, and `No` otherwise.
You can refer to [Lab 1](/labs/lab-1.html#question-6) if you need a refresher on how to do this.
:::

::: render-commit-push
Render, commit, and push your changes to GitHub with the commit message "Added answer for Question 5".

Make sure to commit and push all changed files so that your Git pane is empty afterward.
:::

### Question 6

Return to the following scatter plot of percentage below poverty vs. percentage of people with a college degree, where the color and shape of points are determined by `state` where you were asked to identify at least one county that is a clear outlier by name.

```{r}
#| label: percollege-percbelowpoverty-scatterplot
#| fig-width: 5
#| fig-height: 3
#| echo: false
#| eval: true

ggplot(
  midwest,
  aes(
    x = percollege, y = percbelowpoverty,
    color = state, shape = state
  )
) +
  geom_point() +
  labs(
    title = "Percentage with a college degree versus poverty rate",
    subtitle = "by State in the Midwest",
    x = "% college educated",
    y = "% below poverty",
    color = "State", shape = "State"
  ) +
  annotate(
    geom = "point",
    x = 7.35, y = 48.5, color = "orange", shape = "square open", 
    size = 4.5, stroke = 2
  ) +
  annotate(
    geom = "point",
    x = 45, y = 10.2, color = "red", shape = "square open", 
    size = 25, stroke = 2
  ) +
  theme_minimal()
```

a\.
In a single pipeline, identify the observations marked in the orange square on the upper left corner.
Your answer should be a data frame with four variables: county, state, percentage below poverty, and percentage college educated.

b\.
In a single pipeline, identify the observations marked in the red square in the plot above.
Your answer should again be a data frame with four variables: county, state, percentage below poverty, and percentage college educated.

c\.
Bring your answers from part (a) and part (b) together!
In a single pipeline, and a single `filter()` statement, identify the observations marked in the red and orange square in the plot above.
Your answer should again be a data frame with four variables: county, state, percentage below poverty, and percentage college educated.

d\.
Create a new variable in `midwest` called `potential_outlier`.
This variable should take on the value:

-   `Yes` if the point is one the ones you identified in part (c), i.e., one of the points marked in the squares in the plot above.

-   `No` otherwise.

Then, display the updated `midwest` data frame, with county, state, percentage below poverty, percentage college educated, `potential_outlier` as the selected variables, arranged in ascending order of `potential_outlier`.

c\.
Recreate the visualization above, i.e. a scatterplot of percentage below poverty vs. percentage of people with a college degree, however color the points by the newly created `potential_outlier` variable instead of state.

::: render-commit-push
Render, commit, and push your changes to GitHub with the commit message "Added answer for Question 6".

Make sure to commit and push all changed files so that your Git pane is empty afterward.
:::

### Question 7

a\.
In a single pipeline, calculate the total population for each state and save the resulting data frame as `state_population` and display it in descending order of total population.

b\.
Then, in a separate pipeline, calculate the proportion of the total population in each state and, once again, display the results in descending order of proportion of population.

::: callout-tip
In answering parts (a) and (b), you'll create two new variables, one for total population and other for proportion of total proportion.
Make sure to give them "reasonable" names – short but evocative.
:::

c\.
Which Midwestern state is most populous and what percent of the Midwest population lives there?
Which is the least populous and what percent lives there?

::: render-commit-push
Render, commit, and push your changes to GitHub with the commit message "Added answer for Question 7".

Make sure to commit and push all changed files so that your Git pane is empty afterward.
:::

### Question 8

Calculate the average percentage below poverty for each state and save the resulting data frame as `state_poverty` with the columns `state` and `mean_percbelowpoverty`.

Then, in a new pipeline, display the `state_poverty` data frame in ascending order of `mean_percbelowpoverty`.
Which state has the lowest average percentage below poverty across its counties?
Which state has the highest average percentage below poverty across its counties?

::: render-commit-push
Render, commit, and push your changes to GitHub with the commit message "Added answer for Question 8".

Make sure to commit and push all changed files so that your Git pane is empty afterward.
:::

## Part 2

**All about `group_by()`!**

### Question 9

Suppose we have the following data frame:

```{r}
#| eval: true

df <- tibble(
  var_1 = c(10, 20, 30, 40, 50),
  var_2 = c("Pizza", "Burger", "Pizza", "Pizza", "Burger"),
  var_3 = c("Apple", "Apple", "Pear", "Pear", "Banana")
)

df
```

a\.
What does the following code chunk do?
Run it and analyze the result and articulate in words what `arrange()` does.

```{r}
#| eval: false

df |>
  arrange(var_2)
```

b\.
What does the following code chunk do?
Run it and analyze the result and articulate in words what `group_by()` does.
Also comment on how it's different from the `arrange()` in part (a)?

```{r}
#| eval: false

df |>
  group_by(var_2)
```

c\.
What does the following code chunk do?
Run it and analyze the result and articulate in words what the pipeline does.

```{r}
#| eval: false

df |>
  group_by(var_2) |>
  summarize(mean_var_1 = mean(var_1))
```

d\.
What does the following code chunk do?
Run it and analyze the result and articulate in words what the pipeline does.
Then, comment on what the message says.

```{r}
#| eval: false

df |>
  group_by(var_2, var_3) |>
  summarize(mean_var_1 = mean(var_1))
```

e\.
What does the following code chunk do?
Run it and analyze the result and articulate in words what the pipeline does, especially mentioning what the `.groups` argument does.
How is the output different from the one in part (d).

```{r}
#| eval: false

df |>
  group_by(var_2, var_3) |>
  summarize(mean_var_1 = mean(var_1), .groups = "drop")
```

f\.
What do the following pipelines do?
Run both and analyze their results and articulate in words what each pipeline does.
How are the outputs of the two pipelines different?

```{r}
#| eval: false

df |>
  group_by(var_2, var_3) |>
  summarize(mean_var_1 = mean(var_1), .groups = "drop")

df |>
  group_by(var_2, var_3) |>
  mutate(mean_var_1 = mean(var_1))
```

::: render-commit-push
Render, commit, and push your changes to GitHub with an appropriate commit message.

Make sure to commit and push all changed files so that your Git pane is empty afterward.
:::

### Question 10

Did you select your pages on Gradescope?
You don't need to write an answer for this question, if you select your pages when you upload your lab to Gradescope, you'll get full points on this question.
Otherwise, you'll get a 0 on this question.[^2]

[^2]: We're assigning points to this seemingly trivial task because not selecting your pages and questions will greatly slow down the grading.
    So we want to make sure you're properly motivated to complete this task!

# Wrap-up

## Submission

Once you are finished with the lab, you will submit your final PDF document to Gradescope.

::: callout-warning
Before you wrap up the assignment, make sure all of your documents are updated on your GitHub repo.
We will be checking these to make sure you have been practicing how to commit and push changes.

You must turn in a PDF file to the Gradescope page by the submission deadline to be considered "on time".
:::

To submit your assignment:

-   Go to <http://www.gradescope.com> and click *Log in* in the top right corner.
-   Click *School Credentials* $\rightarrow$ *Duke NetID* and log in using your NetID credentials.
-   Click on your *STA 199* course.
-   Click on the assignment, and you'll be prompted to submit it.
-   Mark all the pages associated with question. All the pages of your lab should be associated with at least one question (i.e., should be "checked").

::: callout-important
## Checklist

Make sure you have:

-   attempted all questions
-   rendered your Quarto document
-   committed and pushed everything to your GitHub repository such that the Git pane in RStudio is empty
-   uploaded your PDF to Gradescope
-   selected pages associated with each question on Gradescope
:::

## Grading

The lab is graded out of a total of 50 points.

On Questions 1 through 9, you can earn up to 5 points on each question:

-   5: Response shows excellent understanding and addresses all or almost all of the rubric items.

-   4: Response shows good understanding and addresses most of the rubric items.

-   3: Response shows understanding and addresses a majority of the rubric items.

-   2: Response shows effort and misses many of the rubric items.

-   1: Response does not show sufficient effort or understanding and/or is largely incomplete.

-   0: No attempt.

On Question 10, you can earn either 0 or 5 points for selecting questions and pages on Gradescope:

-   5: Selected questions and pages on Gradescope.
-   0: Did not select some or all questions and pages on Gradescope.
